---
title: "Problem Set 1 [Applied Stats/Quant Methods 1]"
author: "Emily Zhang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Question 1: Education 

## 1. 
```{r}
y <- c(105, 69, 86, 100, 82, 111, 104, 110, 87, 108, 87, 90, 94, 113, 112, 98, 80, 97, 95, 111, 114, 89, 95, 126, 98)
```

### number of observations 
```{r}
n = length(y)
```

### descriptive statistics
```{r}
mean_y <- sum(y)/n
sd_y <- sqrt(sum((y - mean_y) ^ 2) / (n - 1))
se_y <- sd_y/sqrt(n)
t_score <- qt(0.95, df = n - 1)

ci_90_lower= mean_y - t_score * se_y
ci_90_upper = mean_y + t_score * se_y
```


### hypothesis test 
### variables:
### explanatory: students (numeric)
### response: IQ scores (numeric)

### visualizing the distribution
```{r}
hist(y,
     breaks = 8,
     probability = TRUE,
     main = "IQ Scores of Students",
     xlab = "Scores")

lines(density(y), col = "blue")
```

## 2.

### significance tests
### question: Is the average IQ score in the counselor's school from the sample higher than the average IQ score (100) among all the schools in the country?

### hypotheses: 
# H0: average IQ score of the students is 100                   (mu = 100)
# HA: average IQ score of the students is greater than 100      (mu > 100)
```{r}
t.test(y, mu = 100, alternative = "greater", conf.level = 0.95)
```



# Question 2: Political Economy 
```{r}
expenditure <- read.table("https://raw.githubusercontent.com/ASDS-TCD/StatsI_2026/main/datasets/expenditure.txt", header=T)
```

## 1. 

### quick overview 
```{r}
head(expenditure)
str(expenditure)
summary(expenditure)
```

### loading libraries
```{r}
library(ggplot2)
```

```{r}
vars <- c("Y", "X1", "X2", "X3")

labels <- c(
  "Expenditure",
  "Income",
  "Financially Insecure",
  "Urban"
)

par(mfrow = c(2, 3))

for (i in 1:(length(vars) - 1)) {
  for (j in (i + 1):length(vars)) {
    
    plot(
      expenditure[[vars[j]]],
      expenditure[[vars[i]]],
      xlab = labels[j],
      ylab = labels[i],
      main = paste(labels[i], "vs", labels[j]),
      cex = 0.4
    )
    
    abline(
      lm(expenditure[[vars[i]]] ~ expenditure[[vars[j]]]),
      col = "blue",
    )
    
    r <- cor(
      expenditure[[vars[i]]],
      expenditure[[vars[j]]]
    )
    
    mtext(
      paste("r =", round(r, 2)),

    )
  }
}
```

## 2. 

```{r}
ggplot(expenditure, aes(x = factor(Region), y = Y, color = factor(Region))) +
  geom_boxplot() +
  stat_summary(
    aes(shape = "Mean"),
    fun = mean,
    geom = "point",
    size = 3, 
    alpha = 0.45
  ) +
  labs(
    title = "Housing Expenditure by Region",
    x = "Region",
    y = "Per Capita Expenditure",
    color = "Region",
    shape = ""
  ) +
  scale_shape_manual(
    values = c("Mean" = 18)
  )
```

## 3. 
```{r}
ggplot(expenditure, aes(x = X1, y = Y)) +
  geom_point() +
  labs(
    title = "Housing Expenditure vs Income",
    x = "Per Capita Income",
    y = "Per Capita Expenditure"
  )
```

```{r}
ggplot(
  expenditure,
  aes(
    x = X1,
    y = Y,
    color = factor(Region),
    shape = factor(Region)
  )
) +
  geom_point() +
  labs(
    title = "Housing Expenditure vs Income by Region",
    x = "Per Capita Income",
    y = "Per Capita Expenditure",
    color = "Region",
    shape = "Region"
  )
```

